An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-07-30 EconGrowth.Rnw
Added note about stages of growth.
Added note about possibly eliminating q.

@@

@@

>>>>

Should we eliminate $u$ data? Only 3 countries. Only 20 years. Doesn't add much insight. 
Too bad, but perhaps necessary to simplify paper.

@@

Next steps:

Stages of growth crowd implicitly assumes countries at similar stage of growth will
have similar values for $\alpha$, $\beta$, $\gamma$. However, our results show
that not to be true. Needs further exploration to cover more (all?) countries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Interpretation of Results}
\label{sec:Interpretation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

**** Notes for interpretation of results ****

Country interpretation
Comparison Cobb Douglas and CES results.
US: 
CD: Energy is unnecessary
CES (kl)e: energy is unnec (delta is 1); also (el)k energy is unnec. (delta_1 is 0); (le)k no indication
Relative consistent signal that energy is unnec.
Sigma:
capital and energy are complements
labour and energy are subsitutable

UK:
CD: Capital is unnec.
CES (kl)e: energy is unnec (delat is 1); (le)k capital is unnec. (delta is 1); (ek)l (delta_1 is 0) energy is unnec,
Thus, inconsistent results, possibly look at different time series as massive changes in economic policy in Uk economic history
Sigma:
energy is not substitutable with capital or labour, totally complementary

JP:
CD: energy is unnec.
CES: (kl)e; (le)k; (ek)l: no indications, all are important. weak signal that labour is less necessary (kl(e) delta_1 =0.86 and (ek)l delta =0.89)
Thus, inconsistent results.
Sigma:
labour and energy are complements
capital and energy complements

CN:
CD: energy not necessary, capital less necessary
CES: (kl)e: energy is unnec. (delta is 1); (le)k labour is unnnec (delta_1 is 0); (ek)l energy is unnec. (delta_1 close to 0)
Strong signal that energy is unnec, but the labour and capital signals are conflicting
Sigma:
capital and energy is complementary, but labour and energy more substitutable

ZA:
CD: Labour is unnec.
CES: (kl)e: labour is unnec. (delta_1 is 1); (le)k labour is unnec. (delta_1 is 0); (ek)l labour is unnec. (delta is 1)
Consistent signal that labour is unnec.
Sigma:
capital and energy is complementary
labour and energy is substitutable

SA:
CD: no indications
CES: no indications, but weak signal that labour is unnec in (ek)l (delta is 0.93) 
Almost consistent signal that no indocations.
Sigma:
labour and energy are complements
capital and energy are substitutes

IR:
CD: energy is unnec.
CES: (kl)e: energy is unnec (delta=1); (le)k: energy is unnec. (delta_1=1); (ek)l: labour is unnec. (delta=1)
sigma:
labour and energy still complements, but not as strong
capital and energy complements

TZ:
CD: labour in unnec.
CES: (kl)e: labour is unnec (delta_1=0.97); (le)k: weak signal that energy is unnec. (delta_1=0.92); very weal signal capital is unnec (delta=0.84); (ek)l: labour is unnec. delta=0.96 
Consistent signal that labour is unnec, with exception of one nesting.
Sigma:
labour and energy complements
capital and energy perfect substitutes

ZM:
CD: labour is unnec; energy is unnec.
CES: (kl)e: labour is unnec. (delta_1=1); labour is unnec (le)k: (delta_1=0); (ek)l: labour is unnec. (delta=1)
Consistent signal that labour is unnec.
Sigma:
labour and energy complements, not as strong
capital and energy perfect substitutes.

Discussion of results: How to get better consistent signals?
- Use CD as starting points and do CES convergence on that basis.
- Do scatterplots to get range of uncertainty inputs, then re-interpret. delta against sigma plots, as well as delta_1 and sigma_1 plots.

Thus, to do the fitting well we need to understand:
- nesting
- algorithm
- starting values
esp. for CES. 
(Maybe illustrate with one case study such as China what implications are).

How do we compare witj LINEX results?
CES is a broader specification of CD, LINEX is not. In LINEX no delta's nor sigmas are generated. Only efficiency terms are generated in LINEX, thus in other words economic growth is only possible by increased efficiency. 
The efficiency of using energy is driven by capital stock. The objective/growth engine is to increase capital at a rate that is faster then growth in labour and energy.  Increasing labour relative to energy gives a muted benefit. 
LINEX derived for Germany: only really applicable when energy and capital are complements.

Discussion in relation to other papers. 
Problems with other papers:
- If not giving you all parameters
- Not explain nesting approach 
- No CI on data

Economic interpretation (revisit after scatterplot information)
Developed economies: Capital and energy is complementary (US: labour and energy substitutable)
NICs: Capital and energy complementary, but labour and energy more substitutable
OPEC: Labour and energy complementary (mixed results on capital and energy)
LDCs: Labour and energy complementary, Capital and energy perfect substitutes.
